Improving the templates for multicolor Ramsey numbers and Schur numbers

### 2022-06-25 article_ramsey/template_improvement.tex
definitions

\begin{abstract}
Lower bounds for several multicolor Ramsey numbers as well as Schur numbers have recently been improved using a 
construction based on a particular type of colorings nammed "sf-templates". This article describes a generalization of 
the sf-templates: the \(b\)-templates.
\end{abstract}

\section{Introduction}
This article addresses the edge-coloring of complete graphs in an arbitrary number of colors while limiting the clique 
number in each color. In \cite{RowleyRamsey}, a construction using as a repeating pattern a particular type of 
colorings nammed sf-templates is described and used to improve lower bounds on several Ramsey numbers as well as Schur 
numbers.

Ramsey numbers and Schur numbers are defined in Section \ref{sec:def}. Section \ref{sec:temp} defines the 
\(b\)-templates and describes the corrresponding construction. Section \ref{sec:results} states and proves the main 
results on \(b\)-templates and the construction. Section \ref{sec:bounds} records the new inequalities and some new 
lower bounds.

@@

Pour \(n \in \mathbb{N}\), on note \(K_n\) le graphe complet d'ordre \(n\).

\begin{definition}[Coloriage de Ramsey]
Soit \(n \in \mathbb{N}\), la taille du graphe, et soit \(r \in \mathbb{N}\), le nombre de couleurs. Soit \((k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r\). On appelle coloriage de Ramsey tout coloriage des arêtes de \(K_n\) avec \(r\) couleurs tel que pour toute couleur \(c \in [\![1, r]\!]\), le coloriage ne contient pas de sous-graphe complet \(K_{k_c}\) monochromatique en la couleur \(c\). \(R(k_1, ..., k_r ; n)\) désigne l'un (quelconque) de ces coloriages et l'ensemble de ces coloriages est noté \(\mathcal{R}(k_1, ..., k_r ; n)\). Si tous les \(k_c\) sont égaux à un \(k\), on utilise également les notations \(R_r(k; n)\) et \(\mathcal{R}_r(k; n)\).
\end{definition}

\begin{theorem}[Théorème de Ramsey fini \cite{Ramsey}]
\label{thm:ram}
Soit \(r \in \mathbb{N}\) et soit \((k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r\). Alors il existe un entier \(n \in \mathbb{N}\) tel que \(\mathcal{R}(k_1, ..., k_r ; n) = \varnothing\).
\end{theorem}

Le théorème \ref{thm:ram} justifie la définition suivante.

\begin{definition}[Nombre de Ramsey]
Soit \(r \in \mathbb{N}\) et soit \((k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r\). On définit le nombre de Ramsey \(R(k_1, ..., k_r)\) comme le plus petit entier \(n \in \mathbb{N}\) tel que \(\mathcal{R}(k_1, ..., k_r ; n) = \varnothing\). Si tous les \(k_c\) sont égaux à un \(k\), ce nombre est également noté \(R_r(k)\).
\end{definition}

Contrairement aux nombres de Schur, les nombres de Ramsey désignent le plus petit entier tel qu'il n'existe pas de coloriage, et non pas la taille maximale d'un coloriage.

\begin{definition}[Coloriage de Ramsey linéaire]
Soit \(n \in \mathbb{N}\), et soit \(r \in \mathbb{N}\). Soit \((k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r\). On indexe les sommets de \(K_n\) avec les entiers de \([\![0, n - 1]\!]\). On appelle coloriage de Ramsey linéaire tout \(R(k_1, ..., k_r ; n)\) pour lequel la couleur de l'arête \((u, v)\) dépend uniquement de \(|u - v|\). Un \(L(k_1, ..., k_r ; n - 1)\) désigne l'un (quelconque) de ces coloriages et l'ensemble de ces coloriages est noté \(\mathcal{L}(k_1, ..., k_r ; n - 1)\). Si tous les \(k_c\) sont égaux à un \(k\), on utilise également les notations \(L_r(k; n - 1)\) et \(\mathcal{L}_r(k; n - 1)\).
\end{definition}

Le théorème \ref{thm:ram} justifie la définition suivante.

\begin{definition}[Nombre de Ramsey linéaire]
Soit \(r \in \mathbb{N}\) et soit \((k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r\). On définit le nombre de Ramsey linéaire \(L(k_1, ..., k_r)\) comme le plus petit entier \(n \in \mathbb{N}\) tel que \(\mathcal{L}(k_1, ..., k_r ; n - 1) = \varnothing\). Si tous les \(k_c\) sont égaux à un \(k\), ce nombre est également noté \(L_r(k)\).
\end{definition}

\begin{proposition}
\(\forall r \in \mathbb{N}, (k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r, L(k_1, ..., k_r) \leqslant R(k_1, ..., k_r)\)
\end{proposition}

Un \(L(k_1, ..., k_r ; n)\) induit un coloriage de \([\![1, n - 1]\!]\). Réciproquement, tout coloriage de \([\![1, n - 1]\!]\) correspond à un coloriage linéraire de \(K_n\). Le théorème suivant porte sur le lien entre un coloriage sur \([\![1, n]\!]\) et les nombres de cliques du coloriage linéaire de \(K_{n+1}\) associé.

\begin{theorem}[Lien avec les coloriages sans-sommes \cite{AbbottHanson}]
\label{thm:lien-partitions}
Soit \(r \in \mathbb{N}\) et soit \((k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r\). Soit \(n \in \mathbb{N}^*\) et soit \(\gamma\) un coloriage de \([\![1, n]\!]\) avec \(r\) couleurs. Alors les deux conditions suivantes sont équivalentes :
\begin{enumerate}[(i)]
\item dans le coloriage linéaire de \(K_{n+1}\) associé à \(\gamma\), la couleur \(c\) contient un sous-graphe complet de taille \(k_c\)
\item il existe un sous-ensemble \(S\) de \([\![1, n]\!]\) monochromatique de couleur \(c\) et de cardinal \(k_c - 1\) vérifiant : pour tous \(x, y \in S\) distincts, \(\gamma(|x - y|) = c\) 
\end{enumerate}
\end{theorem}

Dans la suite, on utilise également \(L(k_1, ..., k_r ; n)\) pour désigner un coloriage de \([\![1, n - 1]\!]\)  associé à un \(L(k_1, ..., k_r ; n)\).

Le théorème \ref{thm:lien-partitions} donne le corollaire suivant.

\begin{corollary}[Lien avec les nombres de Schur]
\[\forall n \in \mathbb{N}^*, S(n) = L_n(3) - 2 \leqslant R_n(3) - 2\] 
\end{corollary}

>>>>

\begin{abstract}
Lower bounds for several multicolor Ramsey numbers as well as Schur numbers have recently been improved using a 
pattern based construction with a particular type of coloring nammed "sf-template". This article describes a 
generalization of the sf-templates: the \(b\)-templates.
\end{abstract}

\section{Introduction}
This article addresses the edge-coloring of complete graphs in an arbitrary number of colors while limiting the clique 
number in each color, as well as sum-free colorings of positive integers. A pattern based construction which uses a 
particular type of coloring nammed sf-template was described in \cite{RowleyRamsey}. This construction was used to 
improve lower bounds on several Ramsey numbers as well as Schur numbers. Lower bounds were then improved by describing 
larger templates in \cite{schurboyz} for Schur numbers and in \cite{rowleyramseysat} for multicolor Ramsey numbers.

Ramsey numbers and Schur numbers are defined in Section \ref{sec:def}. Section \ref{sec:temp} defines the 
\(b\)-templates and describes the corrresponding construction. Section \ref{sec:results} states and proves the main 
results on \(b\)-templates and on the construction. Section \ref{sec:bounds} records the new inequalities and some new 
lower bounds.

@@

\begin{definition}[Coloring and partition]
Let \(S\) and \(T\) be two sets. A coloring of \(S\) with colors in \(T\) is an application 
\(\gamma: S \rightarrow T\). The partition associated to the coloring \(\gamma\) is the partition 
\((\gamma^{-1}(t))_{t \in T ~\text{and}~ \gamma^{-1}(\{t\}) \neq \varnothing}\). Conversely, the coloring associated to 
a partition \((A_i)_{i \in I}\) for some \(I \subset T\) is the application \(\gamma: S \rightarrow T\) such that 
\(\forall s \in S, \forall t \in T, \gamma(s) = t \iff s \in A_t\).
\end{definition}

The set of positive natural numbers is denoted by \(\mathbb{N}^* = \mathbb{N} \backslash \{0\}\). Given two integers 
\(a, b \in \mathbb{Z}\), the set of integers \(\{a, a + 1, ..., b - 1, b\}\) is denoted by \([\![a, b]\!]\). For 
\(n \in \mathbb{N}\),  the complete graph of order \(n\) is denoted by \(K_n\).

\begin{definition}[Ramsey coloring]
Let  \(n \in \mathbb{N}\), the order of the graph, and let \(r \in \mathbb{N}\), the number of colors. Let 
\((k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r\). A Ramsey coloring is an edge coloring of \(K_n\) with 
\(r\) colors such that for any color \(c \in [\![1, r]\!]\), the coloring does not contain any monochramatic complete 
sub-graph of order \(k_c\) with color \(c\). The set of these colorings is denoted by 
\(\mathcal{R}(k_1, ..., k_r ; n)\).
\end{definition}

\begin{theorem}[Ramsey's theorem \cite{Ramsey}]
\label{thm:ram}
Let \(r \in \mathbb{N}\) and let \((k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r\). Then there is 
\(n \in \mathbb{N}\) such that \(\mathcal{R}(k_1, ..., k_r ; n) = \varnothing\).
\end{theorem}

Theorem \ref{thm:ram} leads to the following definition.

\begin{definition}[Multicolor Ramsey number]
Let \(r \in \mathbb{N}\) and let \((k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r\). The multicolor Ramsey  
number \(R(k_1, ..., k_r)\) is defined as the smallest integer \(n \in \mathbb{N}\) such 
\(\mathcal{R}(k_1, ..., k_r ; n) = \varnothing\). If all the \(k_c\)'s are equal to some \(k\), this Ramsey number is 
also denoted by \(R_r(k)\).
\end{definition}

The construction described in \cite{RowleyRamsey} as well as the one described in this article use a subset of Ramsey 
colorings, the linear Ramsey colorings.

\begin{definition}[Linear Ramsey coloring]
Let \(n \in \mathbb{N}\), and let \(r \in \mathbb{N}\). Let \((k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r\). 
We assume that the vertices of \(K_{n + 1}\) are the integers from \([\![0, n]\!]\). A linear Ramsey coloring is any 
Ramsey coloring such that the color of the edge \((u, v)\) only depends on the value of \(|u - v|\). The set of these 
colorings is denoted by \(\mathcal{L}(k_1, ..., k_r ; n)\). The integer coloring associated to a linear Ramsey coloring 
is the coloring of \([\![1, n]\!]\) such that every \(x \in [\![1, n]\!]\) is colored with the color of the edges \((u, v)\) 
such that \(|u - v| = x\).
\end{definition}

Unlike Ramsey colorings, the set \(\mathcal{L}(k_1, ..., k_r ; n)\) corresponds to graph colorings of order \(n + 1\). 
This definition is chosen in this article because linear Ramsey colorings are studied using their associated integer 
coloring. Also, in this article the linear Ramsey numbers are defined as the largest size of a coloring (contrary to 
Ramsey numbers) because it is more convenient in inequalities.

\begin{definition}[Linear Ramsey number]
Let \(r \in \mathbb{N}\) and let \((k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r\). The linear Ramsey number 
\(L(k_1, ..., k_r)\) is defined as the largest integer \(n \in \mathbb{N}\) such that 
\(\mathcal{L}(k_1, ..., k_r ; n) \neq \varnothing\). If all the \(k_c\)'s are equal to some \(k\), this linear Ramsey 
number is also denoted by \(L_r(k)\).
\end{definition}

Linear Ramsey numbers can be used to construct lower bounds for Ramsey numbers because a linear Ramsey coloring is 
also a Ramsey coloring. Therefore for all \(r \in \mathbb{N}\) and for all 
\((k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r\), \(R(k_1, ..., k_r) \geqslant L(k_1, ..., k_r) + 2\).

Let \(n \in \mathbb{N}\). Any linear coloring of \(K_{n + 1}\) induces a coloring of \([\![1, n]\!]\). Conversely, any 
coloring of \([\![1, n]\!]\) corresponds to a linear coloring of \(K_{n + 1}\). Theorem \ref{thm:linkintegers} gives a 
link between some sum-free properties of the coloring of \([\![1, n]\!]\) and the clique numbers of the associated 
coloring of \(K_{n+1}\).

\begin{theorem}[Link to sum-free colorings \cite{AbbottHanson}]
\label{thm:linkintegers}
Let \(r \in \mathbb{N}\) and let \((k_c)_{1 \leqslant c \leqslant r} \in {\mathbb{N}^*}^r\). Let \(n \in \mathbb{N}^*\) 
and let \(\gamma\) be a coloring of \([\![1, n]\!]\) with \(r\) colors. Then the two following conditions are equivalent.
\begin{enumerate}[(i)]
\item There is a monochromatic subgraph of order \(k_c\) with color \(c\) in the linear coloring of \(K_{n+1}\) 
	associated  to \(\gamma\).
\item There is a monochromatic subset \(S\) of \([\![1, n]\!]\) with color \(c\) and cardinality \(k_c - 1\) such that 
	for all distinct \(x, y \in S\), \(\gamma(|x - y|) = c\).
\end{enumerate}
\end{theorem}

\begin{definition}[Schur number]
A subset \(A\) of \(\mathbb{N}\) is said to be sum-free if \(\forall (a, b) \in A^2, a + b \notin A\). Let 
\(n \in \mathbb{N}\). Schur proved in \cite{Schur1917} that there is a largest integer denoted by \(S(n)\) such that 
\([\![1, S(n)]\!]\) can be partioned into \(n\) sum-free subsets. \(S(n)\) is called the \(n\)\textsuperscript{th} 
Schur number.
\end{definition}

Theorem \ref{thm:linkintegers} shows a link between Schur numbers and Ramsey numbers:
\[\forall n \in \mathbb{N}^*, S(n) = L_n(3) \leqslant R_n(3) - 2\].
